The Design and Implementation of the Synchronous Language \CEU

### 2016-10-06 paper.tex
(>) Noemi corrections.

\CEU also enforces the use of \code{finalize} for system calls that deal with
pointers representing resources, as illustrated in the two examples of
Figure~\ref{lst.fin.ceu}:
%
\begin{itemize}
\item If \CEU \textbf{passes} a pointer to a system call (ln. \ax:5), the
pointer represents a \textbf{local} resource (ln. \ax:2) that requires finalization
(ln. \ax:7).
\item If \CEU \textbf{receives} a pointer from a system call return (ln. \bx:4),
the pointer represents an \textbf{external} resource (ln. \bx:2) that requires
finalization (ln. \bx:6).
\end{itemize}
%
\CEU tracks the interaction of system calls with pointers and requires 
finalization clauses to accompany them.
%
In the example in Figure~\ref{lst.fin.ceu}.a, the local variable \code{msg} 
(ln. 2) is an internal resource passed as a pointer to \code{\_send\_request} 
(ln. 5), which is an asynchronous call that transmits the buffer in the 
background.
If the block aborts (ln. 11) before receiving an acknowledge from the 
environment (ln. 9), the local \code{msg} goes out of scope and the external 
transmission now holds a \emph{dangling pointer}.
The finalization ensures that the transmission also aborts (ln. 7).
%
In the example in Figure~\ref{lst.fin.ceu}.b, the call to \code{\_fopen} (ln.  
4) returns an external file resource as a pointer.
If the block aborts (ln. 12) during the \code{await A} (ln. 9), the file 
remains open as a \emph{memory leak}.
The finalization ensures that the file closes properly (ln. 6).
%
In both cases, the code does not compile without the \code{finalize}
construct.%
\footnote{
The compiler only enforces the programmer to write the finalization clause, but
cannot check if it actually handles the resource properly.
}

@@

\CEU aims to offer a simpler semantics than Esterel with more determinism and 
fine-grained control over program execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, simplifying the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can figure out which statement executes next following clear linear
rules.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the omnipresence of timers in embedded systems, a dedicated syntax can 
simplify the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}

>>>>

\CEU also enforces the use of \code{finalize} for system calls that deal with
pointers representing resources, as illustrated in the two examples of
Figure~\ref{lst.fin.ceu}:
%
\begin{itemize}
\item If \CEU \textbf{passes} a pointer to a system call (ln. \ax:5), the
pointer represents a \textbf{local} resource (ln. \ax:2) that requires finalization
(ln. \ax:7).
\item If \CEU \textbf{receives} a pointer from a system call return (ln. \bx:4),
the pointer represents an \textbf{external} resource (ln. \bx:2) that requires
finalization (ln. \bx:6).
\end{itemize}
%
\CEU tracks the interaction of system calls with pointers and requires 
finalization clauses to accompany them.
%
In the example in Figure~\ref{lst.fin.ceu}.a, the local variable \code{msg} 
(ln. 2) is an internal resource passed as a pointer to \code{\_send\_request} 
(ln. 5), which is an asynchronous call that transmits the buffer in the 
background.
If the block aborts (ln. 11) before receiving an acknowledge from the 
environment (ln. 9), the local \code{msg} goes out of scope and the external 
transmission now holds a \emph{dangling pointer}.
The finalization ensures that the transmission also aborts (ln. 7).
%
In the example in Figure~\ref{lst.fin.ceu}.b, the call to \code{\_fopen} (ln.  
4) returns an external file resource as a pointer.
If the block aborts (ln. 12) during the \code{await A} (ln. 9), the file 
remains open as a \emph{memory leak}.
The finalization ensures that the file closes properly (ln. 6).
%
In both cases, the code does not compile without the \code{finalize}
construct.%
\footnote{
The compiler only forces the programmer to write the finalization clause, but
cannot check if it actually handles the resource properly.
}

@@

\CEU aims to offer a simpler semantics than Esterel with more determinism and 
fine-grained control over program execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, simplifying the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can figure out which statement executes next following clear
sequential rules.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the omnipresence of timers in embedded systems, a dedicated syntax can 
simplify the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}
